---
title: "index.Rmd"
author: "njkpark"
date: "5/16/2020"
output: html_document
---
```{r, include=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
library(plotly)
source("analysis.R")
```

## Summary Information
In 2018, there were `r total_shootings` shootings that occurred in the United States. In each of these events, there were never any instances where there were no casualties or injuries. In 2018 alone, `r total_lives_lost` lives were lost due to these shootings and `r total_injured` were injured. `r most_city_impact` city was the most impacted in terms of the amount of times there were shootings in this location. `r most_state_impacted` was the most impacted state in 2018. [(Source: gunviolencearchive.org)](https://www.gunviolencearchive.org/reports/mass-shooting)

## Summary Table
```{r aggregate, echo=FALSE}
agg_summary <- shootings_2018 %>%
  group_by(state) %>%
  summarize("Total Killed" = sum(num_killed),
            "Total Injured" = sum(num_injured),
            "Fatality Rate" = sum(num_killed) /
              (sum(num_killed) + sum(num_injured))) 
kable(agg_summary)
```

This table portrays the total amount of people killed and injured by shootings in each State that was affected in the year 2018. Through this visual, Delaware was the most notable state with a 1.0 fatality rate, however it was not the highest death rate. California had a drastic amount of people that were injured and killed in 2018 compared to other states. 

## Description of Particular Event

A particular event that was local to the University of Washington was the shooting in Seattle (Skyway), `r seattle_state`. This occurred on `r seattle_date`. The total impact was `r seattle_impacted` people, where `r seattle_num_injured` were injured and `r seattle_num_killed` were killed. 
[Seattle Times Report](https://www.seattletimes.com/seattle-news/crime/two-people-killed-in-skyway-shooting/)
```{r, echo=FALSE}
seattle_shooting <- shootings_2018 %>%
  filter(city == "Seattle (Skyway)")
kable(seattle_shooting)
```


## Interactive Map
```{r, echo=FALSE}

suppressWarnings(maps <- ggplot() +
  geom_polygon(data = state_shape,
               aes(x = long, y = lat, group = group),
                color = "gray88", fill = "lavenderblush1") +
  geom_point(data = map_df,
             aes(x = long, y = lat, size = num_killed,
                 color = num_injured,
              text = paste("State:", state,
               "<br>City:", city,
               "<br>Number Killed:", num_killed,
               "<br>Number Injured:", num_injured))) +
  scale_colour_distiller(palette = "RdPu") +
  labs(color = "Number Injured",
       x = "Longitude",
       y = "Latitude",
       title = "Shootings in the U.S. During 2018"))
ggplotly(maps, tooltip = "text")
```

From this interactive map, we can determine that over half of shootings within the U.S. occured in the East. Correspondingly, there is virtually no incidences of shootings in Oregon, Idaho, Montana, Wyoming, North and South Dakota. 

## Plot

With this plot, the question that is aimed to be answered is what the fatality rate was for each state during 2018. A column layout was the most valid method to answer this question because we are dealing with one nominal variable (states) and one categorical variable (fatality rate). 
```{r, echo=FALSE}
plot_map <- shootings_2018 %>%
  group_by(state) %>%
  summarize(fatality = sum(num_killed) / (sum(num_killed) + sum(num_injured)))

ggplot(plot_map) +
  geom_col(mapping = aes(x = fatality, y = state), fill = "black") +
  labs(x = "Fatality Rate", y = "State", title = "Fatality Rate in States") +
  coord_cartesian(xlim = c(0, 1))
```

This plot describes what the rate of death (during a shooting event) was in any particular state during 2018. The calculation would be as follows: death/(death + injury) to get the rate. From this plot, Delaware was the only state that had a 100% death rate (where there were 5 deaths and 0 injured). Utah, Massachusetts, and Conneticut had a death rate of 0% (where there were 4 injured in both Utah and Massachusetts, and 8 injured in Conneticut). 